---
title: 'PDF extraction: Comparing extracted result with "Pension Summary and UA" & "OPEB Summary and UA"'
output: 
  html_document: 
    toc: true
    toc_float: true
---

```{r setup, include=FALSE, results = "asis"}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(stringr)
library(tidyr)
library(knitr)
library(writexl)
```
### 1. General note

* Data extracted matches more than 99% of at least 1 of the values in either Actuarial Assumed Rate of Investment Return in "Pension Summary and UA", Actuarial Assumed Rate of Investment Return and Discoun
t Rate in "OPEB Summary and UA" report. 

* Among the 6 cases without matches, 4 are very likely due to wrong entry in Pension and OPEB, i.e. reported figures are not actually in the pdf files.   

* Pair-wise comparison: Data extracted matches about 88% the Actuarial Assumed Rate of Investment Return in "Pension Summary and UA" report, and about 92% that of "OPEB Summary and UA" report. 


### 2. Data extracted 

- For better legibility, one more column is added - 'context' column. This column displays the line that contains the interested information, 2 lines before and 2 lines after that line.

*Note: These lines are from .txt files, which are converted from pdf files, and are not necessarily the lines as displayed in pfd files. The conversion sometimes destroys lines format.*

```{r}
extracted_full <- rio::import(here::here("data", "output1.csv")) %>% rename("name" = file)

extracted <- extracted_full %>% 
  pivot_longer(cols = 2:12,
               names_to = "search_terms", 
               values_to = "extracted_values") %>%
  mutate(name = str_sub(name, start = 11)) %>% 
  mutate(name = str_trim(name, side = "both")) %>% 
  mutate(extracted_values = str_remove_all(extracted_values, "\\['")) %>% 
  mutate(extracted_values = str_remove_all(extracted_values, "\\']")) %>% 
  mutate(extracted_values = str_remove_all(extracted_values, "%")) %>% 
  mutate(extracted_values = str_remove_all(extracted_values, " percent")) %>% 
  mutate(extracted_values = str_remove_all(extracted_values, regex("'.+"))) 

head(extracted)
```

### 3. Pension Summary and UA

```{r}
pension <- rio::import(here::here("data", "Michigan Pension and OPEB Assumption Data 2020.xlsx"), skip = 5, sheet = "Pension Summary and UA") %>% 
  rename("pension_rate" = `Actuarial Assumed Rate of Investment Return`,
         "name" = `Local Government Name`) %>%
  select(name, pension_rate) %>% 
  mutate(pension_rate = pension_rate * 100) %>% 
  mutate(name = str_replace_all(name, "City Of ", ""))

head(pension)
```

### 4. OPEB Summary and UA

```{r}
opeb<- rio::import(here::here("data", "Michigan Pension and OPEB Assumption Data 2020.xlsx"), skip = 5, sheet = "OPEB Summary and UA") %>% 
  rename("opeb_rate" = `Actuarial Assumed Rate of Investment Return`,
         "name" = `Local Government Name`) %>% 
  select(name, opeb_rate, `Discount Rate`) %>% 
  mutate(opeb_rate = opeb_rate * 100) %>% 
  mutate(name = str_replace_all(name, "City Of ", "")) %>% 
  rename("discount_rate_opeb" = `Discount Rate`) %>% 
  mutate(discount_rate_opeb = discount_rate_opeb * 100)

head(opeb)
```

### 5. Joining 3 data sets

- Join on common names
- Drop all rows that OPEB & Pension do not have corresponding values. Eg. Mcmillan Township

```{r}
d <- left_join(extracted, pension) %>% 
  left_join(opeb) %>% 
  mutate(extracted_values = as.numeric(extracted_values)) %>% 
  #select(-c(2,3)) %>% 
  drop_na(c(extracted_values, opeb_rate, pension_rate)) %>% 
  mutate(extracted_values = str_trim(extracted_values)) %>% 
  mutate(pension_rate = str_trim(pension_rate)) %>% 
  mutate(opeb_rate = str_trim(opeb_rate))
head(d)
```

**In the following comparisons: **

- if OPEB rate/ pension rate equals a value captured by a keyword, return 1, else 0

- for each city name, if the OPEB rate/pension rate is captured by any keywords, the average True/False point will be > 0. 

### 6. Comparison I: Data extracted vs. Pension & OPEB

6 cases where no extracted value matches with either pension_rate, opeb_rate, or discount_rate_opeb
```{r}
d %>% 
  drop_na(extracted_values) %>% 
  mutate(equal_any = ifelse(extracted_values == pension_rate |
                            extracted_values == opeb_rate |
                            extracted_values == discount_rate_opeb, 1, 0)) %>% 
  group_by(name) %>% 
  mutate(average_anymatch = mean(equal_any, na.rm = TRUE)) %>% 
  mutate(got_atleast_1inall = ifelse(average_anymatch > 0, 1, 0)) %>% 
  filter(got_atleast_1inall == 0) %>% 
  select(-c(equal_any, average_anymatch, got_atleast_1inall)) %>% unique() -> not_matched_with_any

not_matched_with_any
write_xlsx(not_matched_with_any,"not_matched_with_any.xlsx")

```

*Manual checking suggests that the values reported in opeb or pension are not found in the files:* 

Ann Arbor: can't find 6.9 in pdf file

Marysville: can't find 7.00 in pdf file

White Lake Charter Township: can't find either 7.35 or 7 in pdf file

Fenton: can't find either 7.35 or 3.5 in pdf file


*Remaining 2 files can't be extracted using the current program due to its broken format:* Hamtramck, Crawford County


### 6. Comparison II: Data extracted vs. Pension 

```{r}
d %>% 
  drop_na(extracted_values) %>% 
  mutate(equal_pension = ifelse(pension_rate == extracted_values, 1, 0)) %>% 
  group_by(name) %>% 
  mutate(average = mean(equal_pension, na.rm = TRUE)) %>% 
  mutate(got_atleast_1 = ifelse(average > 0, 1, 0)) %>% 
  select(-c(equal_pension, average, opeb_rate)) %>% 
  unique() -> extracted_vs_pension
```

- Got at least 1 matches with Pension

```{r}
extracted_vs_pension %>% 
  select(-context) %>% 
  filter(got_atleast_1 == 1)
```

- Did not capture a value that matches Pension

```{r}
extracted_vs_pension %>% 
  filter(got_atleast_1 == 0) %>% 
  select(-got_atleast_1) -> not_matched_with_pension

not_matched_with_pension
write_xlsx(not_matched_with_pension, "not_matched_with_pension.xlsx")
```


### 7. Comparison III: Data extracted vs. OPEB


```{r}
d %>% 
  drop_na(extracted_values) %>% 
  mutate(equal_opeb = ifelse(opeb_rate == extracted_values | 
                             discount_rate_opeb == extracted_values, 1, 0)) %>% 
  group_by(name) %>% 
  mutate(average = mean(equal_opeb, na.rm = TRUE)) %>% 
  mutate(got_atleast_1 = ifelse(average > 0, 1, 0)) %>% 
  select(-c(equal_opeb, average, pension_rate)) %>%
  unique() -> extracted_vs_opeb
```

- Got at least 1 matches with OPEB
```{r}
extracted_vs_opeb %>% 
  select(-context) %>% 
  filter(got_atleast_1 == 1) 

```

- Did not capture a value that matches OPEB

```{r}
extracted_vs_opeb %>% 
  filter(got_atleast_1 == 0) %>% 
  select(-c(got_atleast_1)) -> not_matched_with_opeb

not_matched_with_opeb
write_xlsx(not_matched_with_opeb, "not_matched_with_opeb.xlsx")
```

Some random checking against PDF files shows that these OPEB data points are not correct. 

```{r, function to check on cases}
display_context <- function(cityname) {
  df <- d %>% select(name, context, extracted_values, opeb_rate) %>% 
    filter(name == cityname) %>% slice(1:1)
  
  df$context <- kableExtra::cell_spec(
  df$context, 
  extra_css = "display: block; overflow: scroll; height: 5em; width: 20em;"
)
kableExtra::kable_styling(knitr::kable(df, escape = FALSE))
  
}
```

Eg1. "City of Lowell" has value of 7.35 (see "context" below), while reported value in OPEB is 0. 

```{r}
display_context("Lowell")

# To view all result for Lowell
extracted_vs_opeb %>% 
  filter(name == "Lowell")
```

Eg2. City of Bay City has value of 7.35, OPEB: 7.00

```{r}
display_context("Bay City")

```

Eg3. West Branch has value 7.35, OPEB: 4.25

```{r}

display_context("West Branch")
```

Eg4. Manistee has value of 7.60, OPEB: 0.00

```{r}

display_context("Manistee")

```


### 8. Note on OPEB vs Pension

In many cases, the rates in Pension tab and OPEB tab are not equal. For example: 
```{r}
d %>% 
  mutate(opeb_equal_pension = ifelse(pension_rate == opeb_rate, 1, 0)) %>% 
  filter(opeb_equal_pension == 0)
```


